---
title: 'Collapsible Tree Example: R Markdown'
author: "Adeel Khan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(collapsibleTree)
df <- read.csv("GeographyTable.csv")
```

## Overview

collapsibleTree is an R htmlwidget that allows you to create interactive collapsible Reingold–Tilford tree diagram using D3.js. Turn your data frame into a hierarchical visualization without worrying about nested lists or JSON objects!
  
If you're using Shiny, you can bind the most recently clicked node to a Shiny input, allowing for easier interaction with complex nested objects. The input will return a named list containing the most recently selected node, as well as all of its parents. See the Shiny example for more info.

## Data frames, not lists

When working with data in R, it makes sense (at least to me) to represent everything as a data frame. I'm a big fan of [tidy data](https://cran.r-project.org/web/packages/tidyr/vignettes/tidy-data.html), but this structure does not lend itself to easily designing hierarchical networks. 

collapsibleTree uses [data.tree](https://cran.r-project.org/web/packages/data.tree/vignettes/data.tree.html) to handle all of that, freeing you from a lot of recursive list construction.

Here is an example geography dataset from [data.world](https://data.world/glx/geography-table):

```{r summary}
summary(df)
```

## Rendering the plot

With your data frame in hand, and a vector of columns to graph, creating an interactive collapsible tree diagram can be done like so:

```{r plot}
collapsibleTree(
  df,
  hierarchy = c("continent", "type", "country"),
  root = "Geography",
  width = 800
)
```

## Adding Colors

You can add colors to each node in the diagram manually by declaring a `fill` value. Fill is declared as a hierarchy, applied down the tree. In order to get your vectors to the right length for complex trees (this one has 385 nodes), you could use a pattern like this:

```{r plotcolored}
collapsibleTree(
  df,
  hierarchy = c("continent", "type", "country"),
  root = "Geography",
  width = 800,
  fill = c(
    # The root
    "white",
    # Unique continents
    rep("firebrick", length(unique(df$continent))),
    # Unique types per continent
    rep("steelblue", length(unique(paste(df$continent, df$type)))),
    # Unique countries
    rep("green", length(unique(df$country)))
  )
)
```


## Using Gradients

Throw in some gradients if you'd like! Each node can have its own distinct color. The fill order is depends on the order of the data frame.

```{r plotgradient}
collapsibleTree(
  df,
  hierarchy = c("continent", "type", "country"),
  root = "Geography",
  width = 800,
  fill = c(
    # The root
    "white",
    # Unique continents
    RColorBrewer::brewer.pal(length(unique(df$continent)), "Reds"),
    # Unique types per continent
    colorspace::sequential_hcl(length(unique(paste(df$continent, df$type)))),
    # Unique countries
    colorspace::rainbow_hcl(length(unique(df$country)))
  )
)
```
